--- 
title: "Machine Learning Project"
author: "Mingming Li"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site         
---

# Explonatory Analysis
Training Data : train\
Testing Data  : test
```{r, include=FALSE,warning = FALSE} 
library(tidyverse)
library(caret)
library(leaps)
library(glmnet)
library(ggplot2)
library(earth)
library(ROCR)## performance
library(InformationValue) #ROC curve


train<-read.csv("/Users/mingming/Documents/Machine Learning/Homework1_ML/insurance_t.csv")
test<-read.csv("/Users/mingming/Documents/Machine Learning/Homework1_ML/insurance_v.csv")
#names(train)==names(val)
```
## Target Sampling
```{r, warning=FALSE}
unique(train$INS)
sum(train$INS==1)/length(train$INS)
```
## Split Variables
- Categorical variable\
  - level<10\
  - type=character
```{r, warning=FALSE}
col_unique<-lapply(lapply(train,unique),length)
catag_variable<-names(col_unique[col_unique<=10])
chara_type<-lapply(train,typeof)
chara_names<-names(chara_type[chara_type=="character"])
catag_variable<-unique(c(chara_names,catag_variable))
```
- Continuous variable
```{r, warning=FALSE}
conti_variable<-names(train)
conti_variable<-subset(conti_variable,!(conti_variable%in%catag_variable))
```
## Missing Value
1. Finding the variable that has missing value
2. Imputing the missing value in categorical variable as "Missing"
3. Imputing the missing value in continuous variable as median, and adding related Flag
```{r}
#finding variables that have missing value
na_check=colSums(is.na(train))
na_variable=na_check[na_check>0]
na_variable/nrow(train)
na_cata<-names(na_variable)[names(na_variable) %in% catag_variable]
na_conti<-names(na_variable)[names(na_variable) %in% conti_variable]
flag_matrix=matrix(rep(0,nrow(train)*length(na_conti)),nrow=nrow(train),ncol = length(na_conti))
colnames(flag_matrix)=c(paste(na_conti,"_flag",sep=""))

#impute the continuous variable
for (i in seq(1:length(na_conti))){
  mid_point=median(train[,na_conti[i]],na.rm = TRUE)
  flag_matrix[is.na(train[,na_conti[i]]),i]=1
  train[,na_conti[i]]=replace_na(train[,na_conti[i]],mid_point)
}

#impute the categorical variable
for (i in seq(1:length(na_cata))){
  train[,na_cata[i]]=replace_na(train[,na_cata[i]],"missing")
}
```

## Separation Issue
Relationship between any catagorical variable and the target variable
```{r}
for (i in seq(1,length(catag_variable))){
  print(catag_variable[i])
  print(table(train$INS,train[,catag_variable[i]]))
}
```
Imputing the quasic-complete separation, and adding a corresponding flag
```{r}
train$MMCRED_sep_flag=0
train$MMCRED_sep_flag[train$MMCRED>2]=1
train$MMCRED[train$MMCRED>2]="3+"
new_train=cbind(train,flag_matrix)
```
factorize the categorical variable
```{r}
for(i in c(catag_variable,"MMCRED_sep_flag",colnames(flag_matrix))){
  new_train[,i]=as.factor(new_train[,i])
}
```

## Test data and the Final Data
Do the same data posturing on the test data
```{r, warning = FALSE} 

### NA replace with "missing" or median value
## na_variable store all the vaiables with missing value

na_check_train=colSums(is.na(train))
na_variable_train=na_check[na_check_train>0]

na_check=colSums(is.na(test))
na_variable=na_check[na_check>0]

na_variable %in% na_variable_train

na_variable/nrow(test)
na_cata<-names(na_variable)[names(na_variable) %in% catag_variable]
na_conti<-names(na_variable)[names(na_variable) %in% conti_variable]
flag_matrix=matrix(rep(0,nrow(test)*length(na_conti)),nrow=nrow(test),ncol = length(na_conti))
colnames(flag_matrix)=c(paste(na_conti,"_flag",sep=""))

#impute the continuous variable
for (i in seq(1:length(na_conti))){
  mid_point=median(test[,na_conti[i]],na.rm = TRUE)
  flag_matrix[is.na(test[,na_conti[i]]),i]=1
  test[,na_conti[i]]=replace_na(test[,na_conti[i]],mid_point)
}

#impute the categorical variable
for (i in seq(1:length(na_cata))){
  test[,na_cata[i]]=replace_na(test[,na_cata[i]],"missing")
}

test$MMCRED_sep_flag=0
test$MMCRED_sep_flag[test$MMCRED>2]=1
test$MMCRED[test$MMCRED>2]="3+"
new_test=cbind(test,flag_matrix)

#factorize the categorical variable
for(i in c(catag_variable,"MMCRED_sep_flag",colnames(flag_matrix))){
  new_test[,i]=as.factor(new_test[,i])
}
```
Split the training into two pieces, train and val
```{r}
set.seed(123)
index=sample(1:nrow(new_train),0.7*nrow(new_train))
train=new_train[index,]
val=new_train[-(index),]
test=new_test
```

